---
title: "The Impacts of Aquifer Depletion on the Ability of Groundwater Irrigation to Buffer Against Droughts"
author:
  - Taro Mieno^[University of Nebraska Lincoln, tmieno2@unl.edu], Timothy Foster, Shunkei Kakimoto, Ahmed Chennak, Nicholas Brozovic
output:
  officedown::rdocx_document:
    toc: false
    toc_depth: 1
    number_sections: true
    reference_docx: "word_template.docx"
    plots:
      style: Normal
      align: center
      caption:
       style: Image Caption
       pre: "Figure "
       sep: ": "
    tables:
      style: Table
      layout: autofit
      width: 1.0
      caption:
       style: Table Caption
       pre: "Table "
       sep: ": "
bibliography: DRA.bib
abstract: "Abstract: "
---

```{r echo = F}
library(knitr)
library(here)
library(officedown)
library(officer)

opts_chunk$set(
  fig.align = "center",
  fig.retina = 5,
  warning = F,
  message = F,
  cache = T,
  echo = F,
  fig.cap = TRUE
)
```

```{r cache = F}
#--- packages ---#
library(data.table)
library(tidyverse)
library(flextable)
library(stringr)
library(sf)
library(here)
library(tmap)
library(lfe)
library(modelsummary)
library(patchwork)
```

# List of activities done since April

## Groundwater and Saturated Thickness
+ Conduct a literature review
+ Collect gridMET weather data and assign weather variable values to each of the counties
+ Collect daily weather data up to 2099 under various climate change scenarios
+ Improve the modeling of corn and soybean yield response to weather variables: water deficit (reference evapotranspiration less precipitation)
+ Simulate the impact of saturated thickness on corn and soybean yield under various climate change scenarios
+ Start writing a manuscript for journal submission

## Land value

+ Continually negotiated with ERS to get access to the CoreLogic land value data
+ We are at the final stage of agreement. CoreLogic will send us a document, which all the permitted users will sign.

`r run_pagebreak()`

# Main findings with graphics/tables illustrating key points/findings.

# Key Finging 1:

Figure below show the impact of annual water deficit (the sum of reference evapotranspiration less the sum of precipitation from April through September). The greater the value of water deficit, the more severe drought is. However, it also captures the favorable growing conditions like warm weather at the same time. The figure below shows that groundwater provides strong protection against drought. For dryland production, yield starts to decline precipitously once water deficit goes beyond 0 mm (evapotranspiration and precipitation are equated). However, for irrigated production, corn and soy yields increase up to water deficit of 500 mm to 700 mm depending the level of saturated thickness, capturing the benefit of warm weather while negating damage from water deficit. 

```{r}
include_graphics(here("Shared/Results/y_intensive.pdf"))
```

`r run_pagebreak()`

# Key Finging 2:

Saturated thickness has a large impact of the share of irrigated acres for both corn and soybean. As saturated thickness goes down, well yield (the amount of water that can be pumped per a certain period of time, often measured in gallons per minute) also goes down, which then results in a reduction in irrigated acres. Given the dryland production is so much more vulnerable to drought as shown above, saturated thickness has a large impact of "average" corn and soybean yield.

```{r}
include_graphics(here("Shared/Results/y_ir_share.pdf"))
```

`r run_pagebreak()`

# Key Finding 3:

The total impact of saturated thickness on corn is presented in the figure below, combining the two channels of how saturated thickness affects corn and soybean yield. As expected, the value of saturated thickness is greater in drier years. For example, for a year with water deficit of 1000 mm, the average yield at saturated thickness level of 60 and 450 are about 40 and 220, respectively. This clearly shows that maintaining or increasing saturated thickness levels is crucial from both the economic perspective and food security. 

```{r }
include_graphics(here("Shared/Results/y_sat_impact_total.pdf"))
```
